---
title: "Analysis of the short-squeeze of GameStop stock"
author: "Davide Franceschi, Michele Lombardo, Antonio D'Arienzo"
date: "2025-12-07"
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### RANDOM WALKER'S PROJECT ✨

Define which example we are getting:

```{r}
set.seed(12022002) #davide franceschi
exercise=sample(1:3,1)
```

### Exercise 1

Let's assume the following distributions:

$$
X_t \sim WN(0, \sigma^2) \quad \quad Y \sim N(\mu, \sigma_Y^2)
$$

Let's define the new process $Z_t$:

$$
Z_t = X_t + Y
$$

## Point A

**Mean** $E(Z_t)$ **Calculus**

The mean of $Z_t$ is given by the linearity of the expected value
operator:

$$
E(Z_t) = E(X_t) + E(Y) = 0 + \mu = \mu
$$

**Variance** $Var(Z_t)$ **Calculus**

The variance of $Z_t$ (assuming that $X_t$ and $Y$ are independent) is:

$$
Var(Z_t) = Var(X_t) + Var(Y) = \sigma^2 + \sigma_Y^2
$$

**Autocovariance function** $\gamma_h(t, t+h)$ **Calculus**

The Autocovariance function is defined as:

$$
\gamma_h(t, t+h) = E[(Z_t - E(Z_t))(Z_{t+h} - E(Z_{t+h}))]
$$

Substituting $Z_t = X_t + Y$ and $E(Z_t) = \mu$, we obtain:

$$
= E[(X_t + Y - \mu)(X_{t+h} + Y - \mu)]
$$

Computing the product, we have:

$$
= E[X_t X_{t+h} + X_t Y - \mu X_t + Y X_{t+h} + Y^2 - \mu Y - \mu X_{t+h} - \mu Y + \mu^2]
$$

Applying the linearity of expected value and knowing that: \*
$E(X_t) = 0$ (because $X_t \sim WN(0, \sigma^2)$) \* $E(Y) = \mu$
(because $Y \sim N(\mu, \sigma_Y^2)$) \* $X_t$ e $Y$ are independent
variables, therefore $E(X_t Y) = E(X_t) E(Y) = 0 \cdot \mu = 0$

Let's divide the expression in more rows for a better clarification:

$$
= E[X_t X_{t+h}] + E[X_t Y] - \mu E[X_t] + E[Y X_{t+h}] + E[Y^2] - \mu E[Y] - \mu E[X_{t+h}] - \mu E[Y] + E[\mu^2]
$$

Substituting the values:

$$
= E[X_t X_{t+h}] + 0 - \mu (0) + 0 + E[Y^2] - \mu (\mu) - \mu (0) - \mu (\mu) + \mu^2
$$ Even $E[X_t X_{t+h}]=0 \quad \forall h \neq 0$ From which, everything
simplifies to:

$$
= E[Y^2] - 2\mu^2 + \mu^2
$$

**Calculus of** $E[Y^2]$

*Term* $E[Y^2]$: The Variance of Y is $Var(Y) = E(Y^2) - [E(Y)]^2$, we
can write: $$
    E(Y^2) = Var(Y) + [E(Y)]^2 = \sigma_Y^2 + \mu^2
    $$

**Final result for** $\gamma_h$

Substituting $E[Y^2]$ in the simplifies expression:

$$
\gamma_h =(\sigma_Y^2 + \mu^2) - 2\mu^2 + \mu^2
$$

$$
\gamma_h = \sigma_Y^2
$$

## Point B

**Calculus of the autocorrelation coefficient** $\rho_h$

The autocorrelation coefficient is computed as: $$
\rho_h = \frac{\gamma_h}{\gamma_0}
$$

Substituting values $h \neq 0$:

$$
\rho_h = \frac{\sigma_Y^2}{\sigma^2 + \sigma_Y^2}
$$

Dividing numerator and denominator for $\sigma_Y^2$:

$$
\rho_h = \frac{1}{\frac{\sigma^2}{\sigma_Y^2} + 1} = \frac{1}{1 + \frac{\sigma^2}{\sigma_Y^2}} > 0 \quad \forall h \neq 0
$$

**Conclusion:**

The Autocorrelation coefficient $\rho_h$ is a non-vanishing positive
constant for every lag $h \neq 0$. The process $Z_t$ satisfies the 3
conditions (constant mean and variance, Autocovariance that only depends
on the lag h), therefore we confirm that $Z_t=X_t+Y$ is Weakly
Stationary. However the process is not ergodic because $\rho_h$ is
always positive for every $h \neq 0$, i.e., the process has 'infinite'
memory: the initial draw of $Y$ permanently affects the process level,
preventing a single sample path from exploring the entire state space of
the ensemble.

## Point C

```{r}
set.seed(12022002)
sigma2 = 1
mu = 0
theta = 0.9
T = 100
#X_t = WN(0,sigma2)
X_t = rnorm(T, mean=mu,sd=sqrt(sigma2))
Y = rnorm(1,mean=mu,sd=sqrt(sigma2))
Z_t = X_t + Y
Z_t = ts(data=Z_t)
```

plot of $Z_t$:

```{r}
ts.plot(Z_t) #c.1
```

Visually, the given time series appears as a typical stationary process.
The absence of trends or variations in amplitude (volatility) strongly
suggests that the statistical properties (mean and variance) are
invariant with respect to time t.

```{r}
empirical_mean = mean(Z_t) #c.2
mean(Z_t)
theoretical_mean = mu
theoretical_mean
```

The mean of $Z_t$ is $-0.3439143$ for the first sample of
$Y = -0.3439143$ (consider that $X_t$ is a zero-mean $WN$).

```{r}

empirical_variance = var(Z_t)  #c.2
var(Z_t)
theoretical_variance = sigma2 + sigma2
theoretical_variance
```

It should have been $2$ (as 'theoretical_variance' shows), however in
the sample, $Y$ is no more a r.v., so the variance of $Z_t$ will only
depend on the variance of $X_t$ ($WN$ with variance $= 1$).

```{r}
Z_t_matrix = matrix(data=NA, nrow=T,ncol=5) #nina check c.3
Y_vector = rep(NA,5)
for(i in 1:5){
  X_t = rnorm(T,mean=mu,sd=1)
  Y = rnorm(1,mean=mu,sd=1)
  Y_vector[i] = Y
  Z_t_matrix[,i] = X_t + Y
}
par(mfrow = c(2, 3))

for (i in 1:5) {
  plot(Z_t_matrix[, i],
       type = 'l',
       main = paste("Time Series Y=", round(Y_vector[i], digits= 4)),
       xlab = "times",
       ylab = "Z_t",
       col = i)
       abline(Y_vector[i],b=0)
}

par(mfrow = c(1, 1))
```

As displayed in the five realizations, while each $Z_t$ series behaves
locally as a White Noise (fluctuating around a mean), they settle on
different equilibrium levels determined by the realization of the random
variable $Y$. This visually confirms the lack of ergodicity: the time
average of a single path does not converge to the ensemble mean
(theoretical mean), because it is "trapped" by the specific value of $Y$

```{r}
M = 10000 #monte carlo simulation to evaluate theoretical values c.4
ts_matrix = matrix(data=NA,nrow = T,ncol = M)
for(i in 1:M){
  X_t = rnorm(T,mean=mu,sd=1)
  Y = rnorm(1,mean=mu,sd=1)
  ts_matrix[,i]= X_t + Y
}

mean_vector = apply(ts_matrix,1,mean) #for each t
mean(mean_vector)
var_vector = apply(ts_matrix,1,var)
mean(var_vector)

par(mfrow=c(1,2))
plot(mean_vector, type = 'l')
abline(h=mu, col='red')
plot(var_vector, type = 'l')
abline(h=2,col='red')
par(mfrow=c(1,1))
```

The obtained values (mean of both the 'mean_vector' and 'var_vector')
show a strict relation with the theoretical ones, being very close to
them ($0$ and $2$).

```{r}
library(stats) #c.5
acf(Z_t, lag.max = 50)
theoretical_cor = 1/2
```

As we can see, the sample correlation function 'acf($Z_t$)' has almost
all values not significant, so it is failing to capture the theoretical
correlation of 0.5, This discrepancy arises for the same reason above:
the sample ACF is calculated on a single realization where $Y$ is fixed
(acting as a constant intercept), masking the variance contribution of
$Y$ to the total process variance. Consistent estimation of $\rho_h$
requires ergodicity; since this property does not hold, statistics
derived from a single sample are not consistent estimators of the
population parameters, we have to compute the autocorrelation on many
realizations of the series.

```{r}
lags = c(2,13,22,38) #c.5
t = 5
corrs = matrix(data = NA, ncol = M, nrow = 4, dimnames = list(c('lag2','lag13','lag22','lag38')))

for(i in 1:4){
  h = lags[i]
  corrs[i,] = cor(ts_matrix[t,], ts_matrix[t + h,])
}
```

As shown by the correlations, all the values have quite the same value
$0.5$, proving that we need more than one sample in order to check
whether the time series is ergodic or not.

```{r}
#point c.8
#first case
set.seed(12022002)
sigma2 = 1
sigma_y = 3 #play around with values
mu = 0
mu_y = 2 #play around with values
theta = 0.9
T = 100
#X_t = WN(0,sigma2)
X_t1 = rnorm(T, mean=mu,sd=sqrt(sigma2))
Y1 = rnorm(1,mean=mu_y,sd=sqrt(sigma_y))
Z_t1 = X_t1 + Y1
Z_t1 = ts(data=Z_t1)
ts.plot(Z_t1)

empirical_mean1 = mean(Z_t1)
theoretical_mean1 = mu_y
empirical_variance1 = var(Z_t1)
theoretical_variance1 = sigma2 + sigma_y

M = 10000
ts_matrix1 = matrix(data=NA,nrow = T,ncol = M)
for(i in 1:M){
  X_t1 = rnorm(T,mean=mu,sd=1)
  Y1 = rnorm(1,mean=mu_y,sd=sqrt(3))
  ts_matrix1[,i]= X_t1 + Y1
}
mean_vector1 = apply(ts_matrix1,1,mean)
mean(mean_vector1)
var_vector1 = apply(ts_matrix1,1,var)
mean(var_vector1)
par(mfrow=c(1,2))
plot(mean_vector1, type = 'l')
abline(h=mu_y, col='red')
plot(var_vector1, type = 'l')
abline(h=4,col='red')
par(mfrow=c(1,1))

#second case
set.seed(12022002)
sigma2 = 1
sigma_y2 = 4 #play around with values
mu = 0
mu_y2 = 3 #play around with values
theta = 0.9
T = 100
#X_t = WN(0,sigma2)
X_t2 = rnorm(T, mean=mu,sd=sqrt(sigma2))
Y2 = rnorm(1,mean=mu_y2,sd=sqrt(sigma_y2))
Z_t2 = X_t2 + Y2
Z_t2 = ts(data=Z_t2)
ts.plot(Z_t2)

empirical_mean2 = mean(Z_t2)
theoretical_mean2 = mu_y2
empirical_variance2 = var(Z_t2)
theoretical_variance2 = sigma2 + sigma_y2

M = 10000
ts_matrix2 = matrix(data=NA,nrow = T,ncol = M)
for(i in 1:M){
  X_t2 = rnorm(T,mean=mu,sd=1)
  Y2 = rnorm(1,mean=mu_y2,sd=sqrt(4))
  ts_matrix2[,i]= X_t2 + Y2
}
mean_vector2 = apply(ts_matrix2,1,mean)
mean(mean_vector2)
var_vector2 = apply(ts_matrix2,1,var)
mean(var_vector2)
par(mfrow=c(1,2))
plot(mean_vector2, type = 'l')
abline(h=mu_y2, col='red')
plot(var_vector2, type = 'l')
abline(h=5,col='red')
par(mfrow=c(1,1))

```

$Z_{t1}$ and $Z_{t2}$, that are the new time series considering
different values of the unknown parameters, evolve around the value of
the mean of $Y$, proving the behavior that we observed at the beginning.

The values of 'mean_vector1', 'var_vector1', 'mean_vector2' and
'var_vector2' fluctuate around the imposed theoretical values ($2$ and
$4$ for $Z_{t1}$ and $3$ and $5$ for $Z_{t2}$)

### Exercise 2

For the second exercise of our assignment, we decided to select the time
series of GameStop (GME) because, beyond our shared interest in video
games, it represents an exceptionally complex example of financial data,
making it ideal for time series analysis. Specifically, the period of
the 2021 "short squeeze" introduced extreme volatility and anomalies,
which will require us to apply and evaluate robust models such as ARIMA.

```{r}
library(quantmod)
library(forecast)
library(TSstudio)
library(ggplot2)
# Specify start and end dates
start_date = as.Date("2020-11-01")
end_date = as.Date("2022-11-01")
# Load GME data, using adj price to account for splits/dividends
invisible(getSymbols("GME", src = "yahoo", from = start_date, to = end_date))
GME = na.omit(GME$GME.Adjusted)

```

**Data Preparation**

```{r}
frequency(GME)
library(PerformanceAnalytics)
```

Following the Box&Jenkins procedure here we transformed the data, to
reduce variance, and we have also done the first order difference of the
log-prices through the log-returns.

```{r}
log_GME = log(GME) 
log_GME = ts(log_GME, start=c(2020,213),frequency = 252)
frequency(log_GME)
log_returns = CalculateReturns(GME, method = "log") #returns = first difference of log-prices
log_returns = na.omit(log_returns)
```

*Variance inspection*

```{r pressure, echo=FALSE}
ts.plot(GME)
ts.plot(log_GME)
ts.plot(log_returns)
#we will work using the log-prices due to the heteroscedasticity
ts_surface(GME)
hist(log_returns, breaks = 100); abline(v=mean(na.omit(log_GME)), col = 'red')

```

As we can see all the graphs reveals the extreme volatility event of
January 2021. This peak corresponds to the "short squeeze" phenomenom
driven by retail investors, coordinated via the Reddit comunities, who
decided to challenge the hedge-funds holding short positions. The whole
community came to an agreement on buying and holding the stock, forcing
a massive price surge. Since professional investors reasonably had taken
short positions, this caused significant losses: this is the so called
short-squeeze. Following the peak and the initial enthusiasm, the prices
graually decreased but settled at a higher level compared to the
pre-squeeze conditions. The log-returns verify that volatility returned
to more standard clusters after the event.

```{r}
par(mfrow=c(1,2))
autoplot(acf(coredata(log_GME),plot=FALSE)) #no seasonality, ACF decays slowly suggesting non-stationarity
autoplot(pacf(coredata(log_GME),plot=FALSE))## PACF seems to be going to zero after lag 1, suggesting that log_prices behaves like a random walk

par(mfrow=c(1,1))
```

The ACF plot decays slowly without a clear seasonal pattern, which is a
hallmark of non-stationary processes ($I(1)$). The PACF plot sets to 0
after the first lag, suggesting log-GME behaves as a Random Walk

```{r}

library(tseries)
jarque.bera.test(na.omit(log_returns)) #H0 = normality

```

```{r}
qqnorm(log_returns, main = "QQ Plot of Log_GME ")
qqline(log_returns, col = "red", lwd = 2)
```

Here we used the Jaque-Bera test to check for normality. The result
leads us to reject the null hypothesis ($H_0$: Normality). The QQ-plot
visually confirms this: the theoretical quantiles do not align with the
sample quantiles, particularly at the tails ("fat tails"), indicating
that extreme returns occur more frequently than a Normal distribution
would predict.

```{r}
library(tseries)
par(mfrow=c(1,2))
autoplot(acf(coredata(log_returns),plot=FALSE))
autoplot(pacf(coredata(log_returns),plot=FALSE))
par(mfrow=c(1,1))
```

To notice that in both ACF and PACF the first lag is not significant,
suggesting that the present return is not influenced by the previous one
probably, that is due to the bigger peaks on lag 2,3,4. After the lag 4,
both ACF and PACF 'calm down'

```{r}

adf.test(log_returns, alternative = "stationary") #log-prices are stationary, I(1)
kpss.test(log_returns, null = "Trend") # H0: trend stationarity!
adf.test(log_returns, alternative = "explosive")
```

working with a 1-st order difference (log returns) we get stationarity,
motivating an ARIMA rather than ARMA

## Model Selection

Here we are going to try different model and we are going to choose the
one that fits the data better

```{r}
# log_returns is already I(0) --> ARMA(p,q)
# log_prices is I(1) --> ARIMA(p,1,q)
arima414 = arima(log_GME, c(4,1,4))
arima410 = arima(log_GME, c(4,1,0))
arima014 = arima(log_GME, c(0,1,4))
arima111 = arima(log_GME, c(1, 1, 1))
arima210 = arima(log_GME, c(2, 1, 0))
arima211 = arima(log_GME, c(2, 1, 1))
arima212 = arima(log_GME, c(2, 1, 2))
arima510=arima(log_GME,c(5,1,0))
arima515 = arima(log_GME,c(5,1,5))
arima512= arima(log_GME,c(5,1,2))
arima215=arima(log_GME,c(2,1,5))

AIC(arima111, arima210, arima211, arima212,arima414,arima410,arima014,arima510,arima515,arima215)
BIC(arima111, arima210, arima211, arima212,arima414,arima410,arima014,arima510,arima515,arima215)
```

Among all these ARIMA Models we selected the three with the lowest
AIC/BIC. For example we didn't choose ARIMA515 because even though it's
that one with the lowest AIC, the BIC is much higher, since it penalizes
the more complex models. Now we test the significance of the
coefficients of the selected models.

```{r}

library(lmtest)
coeftest(arima510) 
coeftest(arima414)
coeftest(arima410)#the first coef is not significant
#let's try an arima410 with the ar1 coef=0
sub_set_arima410 <- arima(log_GME, order=c(4,1,0), fixed=c(0, NA, NA, NA))
AIC(sub_set_arima410)
BIC(sub_set_arima410)
```

The best model among those, with the lower AIC/BIC, is the ARIMA410.
However, the first AR coefficient was not statistically significant.
Therefore, to address this, we utilized Gemini support to identify, if
possible, a way for removing just the first coefficient from the model.
This led us to discovered that the command arima itself has the argument
"fixed=c(...)" that allows to do so. The resulting Subset-ARIMA410
yields an improved AIC/BIC compared to the full ARIMA410.

```{r}

print(sub_set_arima410)

```

```{r}
# Fit ARIMA (auto selection)
fit = forecast::auto.arima(log_GME)
print(fit)


```

The automatic suggestion gave us the standard ARIMA410. Therefore our
final decision was to use the Sub-ARIMA410 model because the BIC/AIC are
the lowest among all the model tried and all significant coefs. In
addition, it is also a quite simple model, an AR(4) with just three
parameters. It should work better with new data.

## Model checking/Diagnostic

```{r}
res1 = sub_set_arima410$residuals
```

```{r}
plot(res1)
```

NO PATTERN: residual plot (vs time) shows no clear patterns. However
there's a sort of a rectangular band of scatter only after the
short-squeeze.

```{r}
#UNCORELATION: ACF with most of the value within the confidence level
acf(res1, na.action = na.pass)
tsdisplay(res1)
#TESTS for  SERIAL DEPENDENCE
Box.test(res1, lag = 12, type = "Ljung-Box") # H0: no correlation
#ZERO MEAN: residual plot, histogram, empirical mean
```

Indeed the Box-Ljung p-value is to the limit. Probably it's due to the
first part of the series with the enormous peaks.

```{r}
hist(res1, breaks= 100)
```

The residuals are centered on the zero-mean but not normally
distributed.

```{r}
mean(res1)
#NORMAL DISTRIBUTION: histogram, normal Q-Q plot (standardized residuals), statistical test 
# we test for normality and no correlation of the residuals
jarque.bera.test(na.exclude(res1)) # H0: normality

library(rugarch)
library('nortsTest')
arch.test(res1,arch="Lm")

#conclusion the residuals are uncorrelated and with 0 mean, but they are not normally distributed and heteroscedastic (an ARCH model would be more appropriated)
#since the residuals satisfy the essential properties we defenitely choose the sub_set_arima410
```

The residuals are uncorrelated (Ljung-Box test passes for most lags) and
centered at zero. However, they violate the normality assumption and
exhibit heteroscedasticity (significant ARCH test). While the
Subset-ARIMA410 adequately captures the linear dependence (mean
equation), the presence of ARCH effects suggests that the variance is
not constant. We proceed with this model for the mean forecast, but we
must account for these limitations when constructing prediction
intervals.

## Forecasting

```{r}
h = 30
fcast = forecast(sub_set_arima410, h = h)
fcast$mean

alpha = 0.05
fcast = forecast(sub_set_arima410, h = h, level = 1 - alpha)
```

```{r}
plot(fcast, main = "ARIMA forecast for log_GME")
```

Using the fitted Subset-ARIMA410 model, we computed the forecast for the
next $h=30$ days. The solid line in the plot represents our predictor
$\hat{X}_{t+h}$: the conditional expectation of the future log-prices
given the past history ($\hat{X}_{t+h}=E[X_{t+h} | X_t, X_{t-1}...]$).
However log-prices are not stationary and they behave as a Random Walk
($X_t = X_{t-1}+\epsilon_t$),hence
$\hat{X}_{t+1}=E[X_t + \epsilon_{t+1}|F_t] = X_t$ and the same for the
next forecasts of $\hat{X}_{t+h}$. This is because our prediction is a
straight line at the same level of the last observation.

```{r}
h2 = 5
fcast2 =forecast(sub_set_arima410, h = h2)
fcast2$mean
fcast2 = forecast(sub_set_arima410, h = h2, level = 1 - alpha)
```

```{r}
plot(fcast2, main = " ")
```

For a smaller horizons (e.g., $h=1$ to $5$), the confidence intervals
(prediction intervals) are narrow. As $h$ increases to 30, the
prediction intervals widen significantly instead. Remembering that
log_GME is a Random Walk and therefore it is not stationary
($\gamma_h = t\sigma^2$ for $h \geq 0$), this means that the covariance
depends on the time t. The general h-step ahead forecasting of a Random
Walk can also be written as:
$$X_{t+h} = X_t + \sum_{i=1}^{h} \epsilon_{t+i}$$ The general forecast
is:
$$\hat{X}_{t+h}=E[X_{t+h} | X_t, X_{t-1}...]=\hat{X}_{t+h} = E[X_t + \sum \epsilon_{t+i}] = X_t + \sum E[\epsilon_{t+i}]=X_t$$
As written before. The prediction error is
$$e_{t+h} = X_{t+h} - \hat{X}_{t+h} = (X_t + \sum_{i=1}^{h} \epsilon_{t+i}) - X_t = \sum_{i=1}^{h} \epsilon_{t+i}$$

Namely, the prediction error is exactly the sum of the future shocks The
variance of the error is:

$$Var(e_{t+h}) = Var( \sum_{i=1}^{h} \epsilon_{t+i})=Var(\epsilon_{t+1} + \epsilon_{t+2} + ... + \epsilon_{t+h})$$
Since the $\epsilon$ are independent:

$$Var(\epsilon_{t+h})=Var(\epsilon_{t+1}) + Var(\epsilon_{t+2}) + ... + Var(\epsilon_{t+h})$$
Given the assumption of homoscedasticity ($\epsilon$ is a White Noise):
$$Var(\epsilon_{t+h})=\sigma^2 + \sigma^2 + \dots + \sigma^2=h\sigma^2$$
The variance of the forecast error increases linearly with $h$, this
explain why as h increases the "uncertainty cone" widens.

The prediction intervals generated by the forecast() function in R
assumes that the residuals are uncorrelated and normally distributed.
The calculation follows the formula:\
$$\hat{y}_{t+h} \pm z_{\alpha/2} \cdot \hat{\sigma}_h$$

The first benefit of using confidence intervals is the quantification of
uncertainty: they provide a probabilistic range that allows for better
decision-making and risk assessment.

The main harm here lies in the potential violation of the normality
assumption (the Jarque-Bera test gave $p < 0.05$ as result for the
residuals). Since they are not normally distributed the risk can be
underestimated. They exhibit heavy tails. Consequently, the standard
Gaussian intervals produced by R are likely too narrow, underestimating
the probability of extreme price movements (black swan events). The
other principal problem is the homoscedasticity assumed by standard
ARIMA intervals. Since our ARCH test was significant, the model fails to
adapt the interval width to periods of high volatility clustering.

## Residual Bootstrap

```{r}
set.seed(12022002)
fc_boot = forecast(sub_set_arima410, h=h, bootstrap = TRUE, npaths = 5000, level = 1-alpha)
fc_boot$mean
```

```{r}
plot(fc_boot)
lines(fcast$lower, col='red')
lines(fcast$upper, col='red')
#the bootstrap interval is wider than the normal one
#this means that residuals has fatter tails
#they are also skewed: the upper bound is greater than the lower one.
```

Comparing the intervals, we observe that the Bootstrap intervals are
likely wider and asymmetric compared to the Gaussian ones. This confirms
that the Normal assumption underestimates the risk. Specifically, since
GME has 'fat tails' (high kurtosis), the bootstrap procedure, by
sampling real extreme residuals from the past, projects a larger
potential downside/upside than the strict normal curve would allow.

A key issue with the simple residual bootstrap is that it assumes
residuals are i.i.d. (identically distributed). However, our previous
diagnostic tests revealed ARCH effects (heteroscedasticity). This means
the volatility is not constant: some periods are structurally more
volatile than others. By randomly sampling a residual from a 'calm'
period and plugging it into a forecast, or vice-versa, we ignore the
volatility clustering. We are treating all residuals as exchangeable,
which destroys the information regarding the current state of
volatility. A more advanced method (like a GARCH bootstrap or Block
Bootstrap) would be required to preserve the volatility structure.

## Moving Block Bootstrap

```{r}
library(boot)
#the statistic is the mean daily return 
statistic_function = function(ts_data, index){
  return(mean(ts_data[index]))
}
#l=block length (20 is aproximately the number of trading days per month)
#R=replication number
#sim="fixed" block length
mbb=tsboot(log_returns, mean, R=999, l=20, sim = "fixed")
replStats = as.vector(mbb$t)
print(summary(replStats))
print(boot.ci(mbb,type = c("norm","basic","perc")))
```

```{r}
plot(mbb)
```

Standard bootstrapping destroys the serial dependence structure of the
data. For financial time series like GME, where volatility clustering is
present (as shown by our earlier ARCH test), high-volatility days tend
to follow other high-volatility days. The Moving Block Bootstrap (MBB)
addresses this by resampling consecutive blocks of length $l$ (e.g.,
$l=20$). This preserves the local dependency structure (autocorrelation
and ARCH effects) within each block, ensuring that the simulated series
mimics the 'memory' of the real market. To notice that here we used the
log-returns since the log-prices are non-stationary ($I(1)$).
Bootstrapping the log-prices directly would not yield valid statistics
for the drift. Log-returns are stationary ($I(0)$), satisfying the
stability requirement for bootstrap. The resulting 95% confidence
interval is include zero [-0.0045, 0.0141]. This indicates that the mean
daily return is not statistically distinguishable from zero. This
finding supports the Efficient Market Hypothesis and the Random Walk
nature of the stock: while the volatility is high, there is no
statistically significant deterministic trend (upward or downward) that
can be exploited simply by looking at past mean returns.

## Bootstrap Aggregating

Bagging (Bootstrap Aggregating) improves forecast accuracy using the
BLD-MBB algorithm (Bergmeir et al., 2016). We generated $B=1000$
bootstrapped versions of the original GME series. For each version, we
fitted an ETS model and produced a forecast. The final 'Bagged Forecast'
is the average of these 1000 individual forecasts.

```{r}
n_sims = 1000
sims = bld.mbb.bootstrap(log_GME, n_sims)
h = 30
future_log = matrix(0,nrow = n_sims,ncol = h)
for(i in seq(n_sims)){
  future_log[i,] = simulate(ets(sims[[i]], model = "ZZN"), nsim=h)
}

start = tsp(log_GME)[2]+1/252
simfc = structure(list(
  mean = ts(colMeans(future_log), start = start, frequency = 252),
  lower = ts(apply(future_log, 2, quantile, prob=0.025),
             start = start, frequency = 252),
  upper = ts(apply(future_log, 2, quantile, prob=0.975),
             start = start, frequency = 252), level=95),
  class="forecast")

etsfc = forecast(ets(log_GME), h=h,level=95)
```

```{r}
autoplot(log_GME)+
  ggtitle("Daily log_GME") +
  xlab("Year") + ylab("Log Price") +
  autolayer(simfc, series="Simulated") +
  autolayer(etsfc, series="ETS")
```

Comparing the Bagged forecast with the standard single-model forecast,
the Bagged interval is significantly wider than the ETS one. This
happens because the Bagging procedure resamples the actual historical
residuals of GME, which include the extreme volatility observed during
the 'short squeeze' periods of 2021. Since this was a extraordinary
case, two are the possible views:

-   The standard ETS model appears to underestimate the potential tail
    risk, producing a overly confident (narrow) interval based mostly on
    recent trends. Given the extraordinary nature of the Black Swan, it
    might be considered more realistic.

-   However the Bagged forecast can provide a sort of stress test which
    takes count of more unpredictable scenarios. While the probability
    of a repeat squeeze is low, the Bagging interval correctly signals
    that the potential risk remains far higher than what a standard
    Gaussian model (ETS) implies.

Ideally, an investor should view the ETS interval as the probable range
and the Bagging interval as the possible risk landscape in a
worst/best-case scenario.
